### 2018-06-11 Writing/response_to_reviewers.tex

%TG: add a figure for the difference character dependencies?

We've clarified the sentence with an example as follows:
\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines @@@

\item{\textcolor{blue}{"Correlation between characters detected by the software:"\\
I really struggled with this section. It seems like the problem here is still a coding problem - that the authors of these four characters have chosen four highly dependent characters to build the matrix. Even scoring by hand under parsimony would have problems here. I think what you need to decide here is what you want to communicate in this paragraph. Is it that software is not currently equipped to handle this issue?}}

We've added a paragraph highlithing how the problem indeed comes from the fact that current software are not equipped to handle this issue:

\textit{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} more than the other characters since this would be \textit{a priori} the characters that are the most likely to reflect the group's evolutionary history in terms of split by separating the artiodactyles from the cetaceans.
In other words, this is the correlation between characters that depicts the evolutionary history of this group.
However, the other characters depict the same history (i.e. the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.} lines @@@

@@

\item{\textcolor{blue}{"Additionally, we only consider differences for taxa with shared information (i.e. a Gower distance; Gower, 1971)."\\
I'm not sure what this means. Does this mean that you did not include taxa with missing information when calculating per-character distances?}}

We've clarified this sentence (see Reviewer 1, Major suggestion \ref{gower}).

\item{\textcolor{blue}{"This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state"\\
Why is this important? The text doesn't mention testing anything related to orderedness, or specifying anything related to orderedness in MrBayes.}}

We removed the mention of orderdness of the characters since, indeed, it might be more confusing than helpful in this context (we didn't not simulate any ordered characters and the CD metric does not consider characters as ordered - see supplementary material 1). We've changed the sentence to the following:

\textit{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines @@@

@@

\item{\textcolor{blue}{"However, the effect of these sources of correlation was out of the scope of this study and would have required a posteriori changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest."\\
I'm not sure I follow this.}}
\label{dishonest}

We meant that modifying phylogenetic matrices after the phylogenetic inference is a way to test the effect of the sources of correlation but is not at all a good practice in systematics (it can be seen as dishonest to simply remove characters because they ``don't give the right signal''). We've removed the end of this sentence however to avoid any further confusion.

\item{\textcolor{blue}{"However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree. "\\
I understand why you did this, but I think you should also compare normal to true. That will give us some idea of the scope of how close we can get to true, under these conditions, and will help contextualize the results.}}

We've added the comparisons between the ``normal'' and ``randomised'' trees to the ``true'' tree in the supplementary material 3, figures 6 to 8.
We've added a reference to these figures in the main text:

\textit{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 6, 7 and 8.} lines @@@

### 2018-06-11 Analysis/03-EffectCorrelationFigure.Rmd
Added comparisons to the true trees in the supplementaries

@@

@@

>>>>

<!-- For generating the figure for the supplementary-->

```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
pdf("../Writing/Figures/25t_true_tree.pdf", width = 12, height = 8.25)
op <- par(mfrow=c(2,1))
plot.results.single(t25_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (25 taxa)", legend = TRUE, NTS = TRUE)
plot.results.single(t25_list, metric = 4, what = "true",
    main = "", legend = FALSE, NTS = TRUE)
par(op)
# dev.off()
```

@@

```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
pdf("../Writing/Figures/75t_true_tree.pdf", width = 12, height = 8.25)
op <- par(mfrow=c(2,1))
plot.results.single(t75_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (75 taxa)", legend = TRUE, NTS = TRUE)
plot.results.single(t75_list, metric = 4, what = "true",
    main = "", legend = FALSE, NTS = TRUE)
par(op)
# dev.off()
```

@@

```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
pdf("../Writing/Figures/150t_true_tree.pdf", width = 12, height = 8.25)
op <- par(mfrow=c(2,1))
plot.results.single(t150_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
plot.results.single(t150_list, metric = 4, what = "true",
    main = "", legend = FALSE, NTS = TRUE)
par(op)
# dev.off()
```

### 2018-06-11 Writing/draft_highlighted.tex
Added comparisons to the true trees in the supplementaries

In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states tokens will be used in order to infer the tree.
Some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).

@@

We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
This procedure was used to treat all the characters \hl{as} unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
It also greatly improved the speed of our algorithm implementation to compare the characters.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; \hl{as used in} \citealt{Davalos01072014}).
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).

@@

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 1 and 2).
Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
First, in biology, this tree is always unknown.
Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.

>>>>

In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states tokens will be used in order to infer the tree.
\hl{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} more than the other characters since this would be \textit{a priori} the characters that are the most likely to reflect the group's evolutionary history in terms of split by separating the artiodactyles from the cetaceans.
In other words, this is the correlation between characters that depicts the evolutionary history of this group.
However, the other characters depict the same history (i.e. the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.}

This illustrates how some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).

@@

We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
\hl{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in }
\citealt{felsenstein2004inferring}; \hl{as used in} \citealt{Davalos01072014}).
This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).

@@

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 1 and 2).
Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
First, in biology, this tree is always unknown.
Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 6, 7 and 8.}
